Women in Kenya Severely Lack Knowledge on Source of Condom Supply Across All Age Groups And Provinces

### 2022-04-04 outputs/paper/paper.Rmd
updated Abstract section of the paper.

---
title: "Women in Kenya Severely Lack Knowledge on Source of Condom Supply Across All Age Groups And Provinces"
subtitle: "My subtitle if needed"
author: 
  - Tian Yi Zhang
thanks: "Code and data supporting this paper are available at: AOIWDJAWOIDJ"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

Another interesting statistic in the studies mentioned above are the high
usage rates of female hormonal contraceptives. 12.6% of contraceptive methods
used in the United States are of the oral hormonal type [@citeCDC], and around
9% of surveyees in Kenya report using oral contraceptives as well [@citeDHS].
This statistic is perplexing in that the usage rate of hormonal contraceptives
are rather high despite the noticeable side effects:
Various studies have suggested in the past that female contraceptives often
lead to high-risk side effects. For example, the U.S. Department of Health &
Human Services notes that oral contraceptives that rely on hormonal controls
can lead to higher blood pressure and more blood clots [@citeDHH]. As such, one
would wonder why male contraceptives do not see more wide usage despite a
relative lack of harmful side effects in comparison to female hormonal
contraceptives.

This paper attempts to analyze potential reasons for the relative lack of
condom usage as a contraceptive method in Kenya utilizing statistics presented
in the 1998 DHS final report on Kenya [@citeDHS]. Specifically, the paper focuses
on finding a, of lack thereof, relationship between the age, geographical
location, and the knowledge of condom(TODO)

WRITE SOMETHING ABOUT FINDINGS; WRITE SOMETHING ABOUT THE STURCTURE OF THE PAPER

This paper attempts to analyze potential reasons for the relative lack of
condom usage as a contraceptive method in Kenya utilizing statistics presented
in the 1998 DHS final report on Kenya [@citeDHS]. Specifically, the paper focuses
on finding a, of lack thereof, relationship between the age, geographical
location, and the knowledge of condom(TODO)

The remainder of this paper contains analysis on the dataset and the resulting
findings. Section \@ref(data) details the data analysis of this paper. Section
\@ref(raw-data-origin-and-dhs-methodology) discusses the origin of the raw
data used in this paper, and \@ref(dataset-parsing-and-cleaning) details
the extraction process for the dataset. Section \@ref(results) details the
findings from the data analysis, and all limitations and potential extensions
are discussed in Section \@ref(discussion).

@@

The dataset is then cleaned using the `janitor` package 
[@citejani] and the `pointblank` [@citePB] package is used to set up validation
tests for the class and the content of the datasets.
Subsequent graphs and tables shown in this paper also utilize the
`tidyverse` [@citetidy] and `knitr` [@citeknitr] packages.

@@

As demonstrated in Figure \@ref(fig:graph3), a staggeringly high percentage
of women and men with no education have no knowledge on where to obtain
condoms.
The graph also shows that receiving a higher level of
education leads to an overall decrease in condom source ignorance in
both genders. However, the difference in condom source ignorance between
women and men seem to not be correlated to education level, as the difference
in condom source ignorance between genders is lower in both the 'No Education'
and 'Secondary+' levels compared to the 'Primary Incomplete' and
'Primary Complete' levels.

@@

Across all the tables and graphs presented in this paper, it is apparent that
on average, a significantly higher percentage of women have no knowledge
sources for condoms. It is unclear as to why this disparity exists. One possible
reasoning is that condom is a male-contraceptive and thus less advertised to
women.

>>>>

---
title: "Women in Kenya Severely Lack Knowledge on Source of Condom Supply Across All Age Groups And Provinces"
author: 
  - Tian Yi Zhang
thanks: "Code and data supporting this paper are available at: https://github.com/zhan7818/kenya-condom-knowledge"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "The use of contraception is a vital part of modern family planning, yet high risk hormonal contraceptives are curiously more widely used than male contraceptives such as condoms. We obtain data from the 1998 DHS Program report regarding the primary sources for women and men of Kenya to obtain condoms and analyze it using several graphs and tables. The data shows a consistent trend for percentage of women who do not know a source of condom supply to be much higher than men in Kenya regardless of age groups and geographical locations, though higher education levels correlates to increased knowledge in sources for condom supply. Findings suggest an increase in condom advertisement targeted towards women across all age groups and provinces, and an increase in condom advertisement from governement-run healthy services targeted towards poorly educated citizens."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

Another interesting statistic in the studies mentioned above are the high
usage rates of female hormonal contraceptives. 12.6% of contraceptive methods
used in the United States are of the oral hormonal type [@citeCDC], and around
9% of surveyees in Kenya report using oral contraceptives as well [@citeDHS].
This statistic is perplexing in that the usage rate of hormonal contraceptives
are rather high despite the noticeable side effects:
Various studies have suggested in the past that female contraceptives often
lead to high-risk side effects. For example, the U.S. Department of Health &
Human Services notes that oral contraceptives that rely on hormonal controls
can lead to higher blood pressure and more blood clots [@citeDHH]. As such, one
would wonder why male contraceptives do not see more wide usage despite a
relative lack of harmful side effects in comparison to female hormonal
contraceptives. A possible hypothesis is that due to it being a contraceptive
meant to be used by men, it is less advertised to women.

This paper attempts to analyze potential reasons for the relative lack of
condom usage as a contraceptive method in Kenya utilizing statistics presented
in the 1998 DHS final report on Kenya [@citeDHS]. Specifically, the paper focuses
on finding a, of lack thereof, relationship between the age, geographical
location, and the knowledge of sources for obtaining condoms.

The remainder of this paper contains analysis on the dataset and the resulting
findings. Section \@ref(data) details the data analysis of this paper. Section
\@ref(raw-data-origin-and-dhs-methodology) discusses the origin of the raw
data used in this paper, and \@ref(dataset-parsing-and-cleaning) details
the extraction process for the dataset. 
Subsequent graphs and tables shown in this paper utilize the
statistical programming language R [@citeR] and the
`tidyverse` [@citetidy], `ggplot2` [@citeggplot2] and `knitr` 
[@citeknitr] packages. 
Section \@ref(results) details the
results from the data analysis, which includes findings on Kenyan women's 
generally higher lack of knowledge of a viable source for obtaining condoms 
compared to men. All limitations and potential extensions
are discussed in Section \@ref(discussion).

@@

The dataset is then cleaned using the `janitor` [@citejani] and `dplyr` 
[@citedplyr] package. The `pointblank` [@citePB] package is used to set 
up validation tests for the class and the content of the datasets.

@@

As demonstrated in Figure \@ref(fig:graph3), a staggeringly high percentage
of women and men with no education have no knowledge on where to obtain
condoms. 
The graph also shows that receiving a higher level of
education correlates to an overall decrease in lack of knowledge on sources for
condom supply in
both genders. However, the difference in condom source ignorance between
women and men seem to not be correlated to education level, as the difference
in condom source ignorance between genders is lower in both the 'No Education'
and 'Secondary+' levels compared to the 'Primary Incomplete' and
'Primary Complete' levels.

```{r graph4, fig.cap="Percentage of Surveyees Citing Public Sector as Primary Source for Condom versus Education Level", echo = FALSE, fig.width=8, fig.height=4}
kenya_condom_education %>%
  ggplot(aes(x = education, y = public_sector, color = gender)) +
  geom_point(size = 4) +
  theme_minimal() +
  labs(y = "Cited Public Sector as Primary Source (%)",
       x = "Education")
```

Another important thing to note is shown in Figure \@ref(fig:graph4), as a larger
percentage of both men and women in Kenya cite government-funded public services
as their primary source of condom supply as their education levels increase.
This suggests that an increase in public sector advertisement campaign targeted
towards poorly educated citizens could increase usage rate of condoms.

@@

Across all the tables and graphs presented in this paper, it is apparent that
on average, a significantly higher percentage of women have no knowledge
sources for condoms. It is unclear as to why this disparity exists. One possible
reasoning is that condom is a male-contraceptive and thus less advertised to
women. This suggests that creating more condom advertisements targeted at women
in Kenya could potentially lead to an increase in condom usage as contraceptive
method.

## Higher level of education leads to less ignorance on sources of condom supply

Figure \@ref(fig:graph3) shows that higher levels of education seemingly
correlates with a sharp decrease in ignorance of sources of condom supply.
While it may be difficult to increase the country's educational findings and
quality on a national scale, it could be beneficial to set up clinics or
government-sponsored services that inform the less educated population about
sources of condom supply, especially given what is presented in
Figure \@ref(fig:graph4).
